---
title: "Summary July 2018 Pilot"
author: "Ewan Dunbar and Amelia Kimball"
date: "08/08/2018"
output: html_document
---

```{r setup}
`%>%` <- magrittr::`%>%`
`%dopar%` <- foreach::`%dopar%`
```

## Preprocessing data

Load data

```{r load-data, message=FALSE, warning=FALSE}
DATA_FOLDER <-  "data"
STIMULI_FOLDER <-  "stimuli"

subject_info <- readr::read_csv(paste0(DATA_FOLDER, "/", "presurvey_cleaned.csv"))
results_all <- readr::read_csv(paste0(DATA_FOLDER, "/", "Aggregated_Results.csv"))
postsurvey <- readr::read_csv(paste0(DATA_FOLDER, "/", "postsurvey_cleaned.csv"))

item_info <- readr::read_csv(paste0(STIMULI_FOLDER, "/",
                      "item_meta_information_with_distances.csv")) %>%
  dplyr::mutate(tripletid=paste0("triplet_", tripletid))
```

Process attention checks

```{r attention-checks}
#find attention trials, make them into a df called attention checks
results_all <- results_all %>%
  dplyr::mutate(paid_attention=ifelse(
    grepl('^atten', tripletid),
      ifelse(
        (tripletid=="attention_check_English_F_normalized" | 
         tripletid=="attention_check_francais_F_normalise") & 
        first_sound==1,
        'pass_f', 
        ifelse(
          (tripletid=="attention_check_english_J_normalise" | 
           tripletid=="attention_check_francais_J_normalise") & 
          second_sound==1,
          'pass_j',
          'fail')),
      NA))
```

Filter subjects based on attention checks and surveys

```{r filtering-subjects}
attention_fails <- results_all %>%
  dplyr::filter(!is.na(paid_attention)) %>%
  dplyr::group_by(subject_id) %>%
  dplyr::summarize(N_not_paid_attention=sum(paid_attention == "fail"))

# Only subjects who are in the postsurvey file, i.e., have finished
filtered_subjects <- dplyr::left_join(postsurvey, subject_info,
                                      by =c('subject_id', 'subject_language')) %>%
# Filter based on native language questions
#
# Current logic: subjects answered whether they know any other languages,
# for various different senses of "know", and we filter out based on these
# questions.
#
# This survey is to be replaced by a set of binary questions,
# including (1) early exposure to the "target" language,
# (2) early exposure to other languages; (3) late exposure to
# other languages above a certain amount and/or before a certain
# age and/or with a certain level of fluency

# Remove subjects whose response to 'languages between 0 and 3'
# is not either 'English' or '1' or 'French'
  dplyr::filter(toupper(`lang_0-3`) %in% c("ENGLISH", "1.0", "1",
                                         "FRANÇAIS")) %>%
# Phonetic training filtering - exclude anyone with any classes
# phonet/phonol/linguistics
  dplyr::filter(phonet_class_Y == 0,
                phonog_class_Y == 0,
                ling_course_Y == 0) %>%
# Exclude speech/hearing/vision problems
  dplyr::filter(hear_vis_Y == 0,
                speech_prob_Y == 0) %>%
# Allow no more than one attention check failure
  dplyr::left_join(attention_fails, by="subject_id") %>%
  dplyr::filter(N_not_paid_attention <= 1)

```

Drop filtered subjects, attention checks and practice trials, include item information,
add columns to determine whether user response was correct, add transformed
comparative acoustic distance columns.

```{r select-items}
results <- dplyr::left_join(filtered_subjects, results_all,
                            by=c('subject_id', 'subject_language')) %>%
  dplyr::left_join(item_info, by='tripletid') %>%
  dplyr::filter(grepl("^triplet_", tripletid)) %>%
  dplyr::mutate(user_resp=factor(ifelse(first_sound == "1", "A", "B")),
                user_corr=as.integer(
                  substr(presentation_order, 1, 1) == user_resp))
distance_types <- names(results) %>%
  (function(x) x[grepl("^distance_", x)]) %>%
  (function(x) sub("_TGT$", "", x)) %>%
  (function(x) sub("_OTH$", "", x)) %>%
  unique
for (dt in distance_types) {
  results[[paste0("difference_log_", dt)]] <-
    log(results[[paste0(dt, "_TGT")]]) - 
      log(results[[paste0(dt, "_OTH")]])
}
distance_difference_colnames <- names(results)[
  grep("^difference_log", names(results))]
pcs <- prcomp(results[,distance_difference_colnames],
              scale=TRUE)$x 
colnames(pcs) <- paste0(
  "difference_log_distance_PC", 1:ncol(pcs)
)
results <- cbind(results, pcs)
```

## Testing various models

**First question: what is the right way to deal with acoustic distance?**

Here our goal is to find a simple model that captures a lot of the variance due to acoustic distance, even when extrapolating, or using a small number of subjects. It seems fine to just have a look at what the accuracy ~ distance plot looks like.

**Start here.** It turns out that extrapolating is really hard, and, here, extrapolation-based statistics always turn out to prefer rather poor predictors. I suggest leave-one-out or an approximation, or some better predictor (perhaps training an unsupervised multi-lingual speech embedding?), to proceed with this question. The consonant data from the second pilotwill be useful 

```{r accuracy-by-distance}
results_by_item <- results %>%
  dplyr::group_by_at(
    dplyr::vars(
      "tripletid", "subject_language",
      tidyselect::starts_with("difference_log"))) %>%
  dplyr::summarize(accuracy=mean(user_corr)) %>%
  dplyr::ungroup()
distance_difference_colnames <- names(results_by_item)[
  grep("^difference_log", names(results_by_item))]
distance_plots <- tibble::tibble(distance=distance_difference_colnames) %>%
  dplyr::mutate(plot=purrr::map(.$distance,
                  function(var_name)
                    ggplot2::ggplot(results_by_item,
                      ggplot2::aes_string(
                        y="accuracy", x=var_name)) +
                    ggplot2::geom_point() +
                    ggplot2::facet_wrap(~ subject_language) +
                    ggplot2::coord_cartesian(ylim=c(0,1)) +
                    ggplot2::geom_smooth()))
print(distance_plots$plot)
```

We assume that this simple model (subject response correct ~ difference in log acoustic distance) will continue to do be appropriate when we move to a larger model that includes other predictors.

First hyperparameter - type of acoustic distance
Second hyperparameter - form of relation
Third hyperparameter - N (do stratified sampling)

```{r TESTING-GROUND}
hadley_update <- function(mod, formula = NULL, data = NULL) {
  call <- getCall(mod)
  if (is.null(call)) {
    stop("Model object does not support updating (no call)", call. = FALSE)
  }
  term <- terms(mod)
  if (is.null(term)) {
    stop("Model object does not support updating (no terms)", call. = FALSE)
  }

  if (!is.null(data)) call$data <- data
  if (!is.null(formula)) call$formula <- update.formula(call$formula, formula)
  env <- attr(term, ".Environment")

  eval(call, env, parent.frame())
}

llik_logistic <- function(model, newdata) {
  yhat <- predict(model, newdata=newdata, type="response")
  y <- newdata[[all.vars(formula(model))[1]]]
  if (!is.numeric(y)) {
    y <- contrasts(factor(y))[as.character(y)] # FIXME - is this what glm does?
  }
  return(sum(y*log(yhat) + (1-y)*log(1-yhat), na.rm=TRUE))
}

refit_without <- function(fit, data, indices) {
   d_left_out <- data[indices,]
   d_remaining <- data[-indices,]
   return(hadley_update(fit, data=d_remaining))
}

loo_vanilla <- function(fit,  data) {
  #loo_points <- 1:nrow(data)
  loo_points <- sample(1:nrow(data), 1000)
  lliks <- foreach::foreach(i=loo_points, .combine=c) %dopar%
      llik_logistic(refit_without(fit, data, i), data[i,]) # FIXME
  return(mean(lliks, na.rm=TRUE))
}

extrapolate <- function(fit, data, proportion, balanced=TRUE) {
  # Find extreme values of X - FIXME should support multiple variables
  x <- data[[all.vars(formula(fit))[2]]]
  ranks <- rank(x) # FIXME - would need to apply PCA here
  sorted_unique_ranks <- sort(unique(ranks))
  n <- as.integer(proportion*length(sorted_unique_ranks)) # FIXME - precond. prop.
  if (balanced) {
    n_top = as.integer(n/2)
    n_bottom = n - n_top
    # FIXME - at least document weird interpretation of proportion
    top_ranks <- sorted_unique_ranks[seq_len(n_top) + length(sorted_unique_ranks) - n_top]
    bottom_ranks <- sorted_unique_ranks[seq_len(n_bottom)]
    indices <- which(ranks %in% c(top_ranks, bottom_ranks))
    return(#tibble::tibble(score=
           llik_logistic(refit_without(fit, data, indices), # FIXME
                         data[indices,]) #, fit_wo_edges=refit_without(fit, data, indices
                #)
    ) 
  } else {
    top_ranks <- sorted_unique_ranks[seq_len(n) + length(sorted_unique_ranks) - n]
    top_indices <- which(ranks %in% top_ranks)
    llik_without_top <- llik_logistic(refit_without(fit, data, top_indices),
                                      data[top_indices,])
    bot_ranks <- sorted_unique_ranks[seq_len(n)]
    bot_indices <- which(ranks %in% bot_ranks)
    llik_without_bot <- llik_logistic(refit_without(fit, data, bot_indices),
                                      data[bot_indices,])
    return(#list(score=
                  mean(llik_without_top, llik_without_bot)
                #,
                #fit_wo_top=refit_without(fit, data, top_indices),
                #fit_wo_bot=refit_without(fit, data, bot_indices)
                #)
    )
  }
}

FORMULAS <- list(
  NFP="user_corr ~ difference_log_distance_normed_filterbank__dtw_pathlength",
  NFS="user_corr ~ difference_log_distance_normed_filterbank__dtw_sum",
  NMP="user_corr ~ difference_log_distance_normed_mfcc__dtw_pathlength",
  NMS="user_corr ~ difference_log_distance_normed_mfcc__dtw_sum",
  UFP="user_corr ~ difference_log_distance_unnormed_filterbank__dtw_pathlength",
  UFS="user_corr ~ difference_log_distance_unnormed_filterbank__dtw_sum",
  UMP="user_corr ~ difference_log_distance_unnormed_mfcc__dtw_pathlength",
  UMS="user_corr ~ difference_log_distance_unnormed_mfcc__dtw_sum",
  PC1="user_corr ~ difference_log_distance_PC1",
  PC2="user_corr ~ difference_log_distance_PC2",
  PC3="user_corr ~ difference_log_distance_PC3",
  PC4="user_corr ~ difference_log_distance_PC4",
  PC5="user_corr ~ difference_log_distance_PC5",
  PC6="user_corr ~ difference_log_distance_PC6",
  PC7="user_corr ~ difference_log_distance_PC7",
  PC8="user_corr ~ difference_log_distance_PC8",
  RAND="user_corr ~ fake_difference_log_distance_random",
  NFP2="user_corr ~ difference_log_distance_normed_filterbank__dtw_pathlength + I(difference_log_distance_normed_filterbank__dtw_pathlength^2)",
  NFS2="user_corr ~ difference_log_distance_normed_filterbank__dtw_sum + I(difference_log_distance_normed_filterbank__dtw_sum^2)",
  NMP2="user_corr ~ difference_log_distance_normed_mfcc__dtw_pathlength + I(difference_log_distance_normed_mfcc__dtw_pathlength^2)",
  NMS2="user_corr ~ difference_log_distance_normed_mfcc__dtw_sum + I(difference_log_distance_normed_mfcc__dtw_sum^2)",
  UFP2="user_corr ~ difference_log_distance_unnormed_filterbank__dtw_pathlength + I(difference_log_distance_unnormed_filterbank__dtw_pathlength^2)",
  UFS2="user_corr ~ difference_log_distance_unnormed_filterbank__dtw_sum + I(difference_log_distance_unnormed_filterbank__dtw_sum^2)",
  UMP2="user_corr ~ difference_log_distance_unnormed_mfcc__dtw_pathlength + I(difference_log_distance_unnormed_mfcc__dtw_pathlength^2)",
  UMS2="user_corr ~ difference_log_distance_unnormed_mfcc__dtw_sum + I(difference_log_distance_unnormed_mfcc__dtw_sum^2)",
  RAND2="user_corr ~ fake_difference_log_distance_random + I(fake_difference_log_distance_random^2)"
)
prepare_model <- function(distance_predictors, d) {
  spec <- FORMULAS[[distance_predictors]]
  spec_and_data <- list(spec=spec, data_indices=1:nrow(d))
  return(spec_and_data)
}

fit_model <- function(spec_name, data_indices, data) {
  d <- data[data_indices,]
  f <- as.formula(FORMULAS[[spec_name]])
  fit <- glm(formula=f, data=d, family="binomial") # FIXME
  stats <- tibble::tibble(
    loo_vanilla=loo_vanilla(fit, d),
    extrapolate_10_balanced=extrapolate(fit, d, 0.1, balanced=TRUE),
    extrapolate_10_unbalanced=extrapolate(fit, d, 0.1, balanced=FALSE),
    extrapolate_half_balanced=extrapolate(fit, d, 0.5, balanced=TRUE),
    extrapolate_half_unbalanced=extrapolate(fit, d, 0.5, balanced=FALSE)
  )
  fit_and_stats <- list(fit=fit, stats=stats)
  return(fit_and_stats)
}

results_for_testing_acoustic_distances <- results
set.seed(100)
results_for_testing_acoustic_distances$fake_difference_log_distance_random <- 
  sample(results_for_testing_acoustic_distances$difference_log_distance_PC8) +
  rnorm(nrow(results), 0, sd(results$difference_log_distance_PC8)/3)
set.seed(NULL)


doParallel::registerDoParallel(cores=6)
model_info <- tibble::tibble(
  M_distance_predictors=names(FORMULAS),
  data_indices=vector("list", length(FORMULAS)),
  fit=vector("list", length(FORMULAS)),
  stats=vector("list", length(FORMULAS))
  )
for (i in 1:nrow(model_info)) {
  spec_and_data_i <- prepare_model(model_info[["M_distance_predictors"]][i],
                                   results_for_testing_acoustic_distances)
  model_info[["data_indices"]][[i]] <- spec_and_data_i[["data_indices"]]
  fit_and_stats_i <- fit_model(model_info[["M_distance_predictors"]][[i]], 
                               spec_and_data_i[["data_indices"]],
                               results_for_testing_acoustic_distances)
  model_info[["fit"]][[i]] <- fit_and_stats_i[["fit"]]
  model_info[["stats"]][[i]] <- fit_and_stats_i[["stats"]]
}
model_info_simple <- model_info %>%
  dplyr::select(-fit, -data_indices) %>%
  tidyr::unnest()

```

```{r plot-model-results}
plot(extrapolate_half_balanced ~ extrapolate_10_unbalanced,
     type="n", data=model_info_simple)
text(extrapolate_half_balanced ~ extrapolate_10_unbalanced, 
     labels=M_distance_predictors, data=model_info_simple)
plot(extrapolate_half_balanced ~ extrapolate_10_balanced,
     type="n", data=model_info_simple)
text(extrapolate_half_balanced ~ extrapolate_10_balanced, 
     labels=M_distance_predictors, data=model_info_simple)
plot(extrapolate_half_unbalanced ~ extrapolate_10_unbalanced,
     type="n", data=model_info_simple)
text(extrapolate_half_unbalanced ~ extrapolate_10_unbalanced, 
     labels=M_distance_predictors, data=model_info_simple)
plot(extrapolate_half_unbalanced ~ extrapolate_10_balanced,
     type="n", data=model_info_simple)
text(extrapolate_half_unbalanced ~ extrapolate_10_balanced, 
     labels=M_distance_predictors, data=model_info_simple)
plot(extrapolate_half_unbalanced ~ extrapolate_half_balanced,
     type="n", data=model_info_simple)
text(extrapolate_half_unbalanced ~ extrapolate_half_balanced, 
     labels=M_distance_predictors, data=model_info_simple)
plot(extrapolate_10_unbalanced ~ extrapolate_10_balanced,
     type="n", data=model_info_simple)
text(extrapolate_10_unbalanced ~ extrapolate_10_balanced, 
     labels=M_distance_predictors, data=model_info_simple)
plot(loo_vanilla ~ extrapolate_half_balanced,
     type="n", data=model_info_simple)
text(loo_vanilla ~ extrapolate_half_balanced, 
     labels=M_distance_predictors, data=model_info_simple)
plot(loo_vanilla ~ extrapolate_10_balanced,
     type="n", data=model_info_simple)
text(loo_vanilla ~ extrapolate_10_balanced, 
     labels=M_distance_predictors, data=model_info_simple)
```

